Legacy Data Confounds Modern Genomics Studies

### 2019-03-18 PlottingScripts/ProcessPlotData.Rmd
update

#Joint Frequency Spectrum
```{r}
if(!file.exists(paste0(DataPath,
                       'jointAF.txt'))) {
  #Load Allele Frequency files
  AF.1kGP <- paste0(DataPath,
                    '1kGP_GenomeWide_JPT_1.frq')
  JPT <- fread(
    AF.1kGP,
    fill = TRUE,
    col.names = c('Chr', 'Pos', 'N_ALLELES',
                  'N_CHR', 'JPT_AF', 'JPT_MAF')
  )
  AF.NAG <- paste0(DataPath,
                   'NAG_GenomeWide.4bed_filtered.freq.frq')
  NAG <- fread(
    AF.NAG,
    fill = TRUE,
    col.names = c('Chr', 'Pos', 'N_ALLELES',
                  'N_CHR', 'NAG_AF', 'NAG_MAF')
  )

  #merge them together to get joint frequency spectrum
  join <- merge(JPT[which(JPT$JPT_AF <= 1), ],
                NAG[which(NAG$NAG_AF <= 1), ],
                by = c('Chr', 'Pos'), all = TRUE)
  join[is.na(join)] <- 0 #missing positions are frequency of 0

@@

  hwe <- merge(NAG_hwe[, c('V1', 'V2')],
               JPT_hwe[, c('V1', 'V2')],
               by = c('V1', 'V2'),
               all = TRUE)
  #remove sites from AF that don't meet HWE
  join <- join[!paste(join$CHROM, join$POS)
               %in% c(paste(hwe$V1, hwe$V2)),]

  join <- join[, c('Chr', 'Pos', 'JPT_MAF',
                   'JPT_AF', 'NAG_MAF', 'NAG_AF')]
  #make numeric
  join <- sapply(join, as.numeric)
  join <- join[complete.cases(join), ]

@@

  #save to file
  write.table(join,
              paste0(DataPath,
                     'jointAF.txt'))
}
```

@@

#get positions of random SNPs as control
set.seed(123)
random.pos <- NotSig[sample(nrow(NotSig), 
                            nrow(sig.6)),
                     c('Chr','Pos')]
write.table(random.pos, 
            file = paste0(DataPath,
                        'random.pos.txt'), 
            col.names = FALSE, 
            row.names = FALSE, 
            quote = FALSE, 
            sep = '\t')
rm(NotSig) #save memory
```
extract context for snps of interest
```{bash}
#Using GenomeWide.AncestralContext.txt
#To get mutation context, use $path/AncestralContext.py
#compare first two columns in both files, extract matches
path='/Users/luke/Documents/QualityPaper/Misc/'
if [ ! -f $path/random.pos.context.txt ]; then
  echo "random pos context"
  awk 'NR==FNR{a[$1,$2];next} ($1,$2) in a' \
  $path/random.pos.txt \
  $path/GenomeWide.AncestralContext.txt \
  > $path/random.pos.context.txt
fi

@@

```{r}
BasPath = '/Users/luke/genomes/bas_file/1000GenomesBAS/'

@@

metaData
#save to file
saveRDS(metaData,
            paste0(DataPath,
                   'metaData.rds'))
```

```{r}
phase1 <- c('YRI','LWK','CHB','CHD','JPT','CEU','LWK')
#read single population regressions
Reg <- fread(paste0(DataPath,
                    'AllRegressionsOver10.txt'), 
             col.names = c('Chr','Pos',
                           'Population','deviance'), 
             colClasses = c('numeric', 'numeric', 
                            'character', 'numeric'))
#Get P-Values and significant sites
Reg$plog10 <- -log10(pchisq(Reg$deviance, 1, 
                            lower.tail = FALSE))
Reg<-Reg[which(Reg$plog10 >= 6),]
#Count number of populations where a variant is significant
NumPop <- as.data.frame(table(Reg$Chr,Reg$Pos))
NumPop <- NumPop[which(NumPop$Freq>1),]
names(NumPop) <- c('Chr','Pos', 'SigHits')
NumPop$Chr<-as.numeric(as.character(NumPop$Chr)) 
NumPop$Pos<-as.numeric(as.character(NumPop$Pos))

@@

#save to file
saveRDS(overLap,
            paste0(DataPath,
                   'overLap.rds'))
```
```{r}
#load 1kgp 
kgp1= paste0(DataPath,
             '1kGP_genomeWide_SINGLEPOP_83.genotypes.txt')
kgp1.n=fread(paste0(DataPath,
                    '1kGP_genomeWide_SINGLEPOP_83.header'))
gt1<-unique(fread(kgp1,
                  col.names = colnames(kgp1.n)))
gt1.m<-melt(gt1, 
            id.vars = c('#CHROM','POS','ID'))
colnames(gt1.m) = c('CHROM','POS',
                    'ID','sample','kGT')

@@

```{r}
#read Genotypes
han1=fread(paste0(DataPath,
                  'SignificantSNPs_Han90_0.01_genotype.txt'))
HanNames=fread(paste0(DataPath,
                      'SignificantSNPs_Han90_COLNAMES.txt'),
               header=F)
names(han1) = as.character(unlist(HanNames))
#melt to long format
han = melt(han1, id=c('#CHROM','POS','ID'))
#fix chromosome names
han$CHROM = as.numeric(as.character(substring(han$'#CHROM',4)))
han$'#CHROM' <- NULL

@@

#save to file
saveRDS(INDEL,
            paste0(DataPath,
                   'INDEL_HanKgp.rds'))
#save to file
saveRDS(SNP,
            paste0(DataPath,
                   'SNP_HanKgp.rds'))
```

>>>>

##GWAS
```{r}
#Load single population regression results
Reg <- fread(
  paste0(DataPath,
         '1KGP.GenomeWide.Regression_JPT.csv'),
  col.names = c('Chr', 'Pos',
                'Population', 'deviance'),
  colClasses = c('numeric', 'numeric',
                 'character', 'numeric'))

#Compute P values from deviance from null model
Reg$plog10 <- -log10(pchisq(Reg$deviance,
                            df = 1,
                            lower.tail = FALSE))

#split data into sig / not sig for plotting
SigPos <- Reg[which(Reg$plog10 > 6), ]
NotSig <- Reg[which(Reg$plog10 < 6), ]
rm(Reg) #save memory

#save to file
saveRDS(SigPos,
        paste0(DataPath,
               'SigPos.rds'))
#save to file
saveRDS(NotSig,
        paste0(DataPath,
               'NotSig.rds'))
```
##Joint Frequency Spectrum
```{r}
if(!file.exists(paste0(DataPath,
               'jointAF.rds'))){

#Load Allele Frequency files
AF.1kGP <- paste0(DataPath,
                  '1kGP_GenomeWide_JPT_1.frq')
JPT <- fread(
  AF.1kGP,
  fill = TRUE,
  col.names = c('Chr', 'Pos', 'N_ALLELES',
                'N_CHR', 'JPT_AF', 'JPT_MAF'))
AF.NAG <- paste0(DataPath,
                 'NAG_GenomeWide.4bed_filtered.freq.frq')
NAG <- fread(
  AF.NAG,
  fill = TRUE,
  col.names = c('Chr', 'Pos', 'N_ALLELES',
                'N_CHR', 'NAG_AF', 'NAG_MAF'))

#merge them together to get joint frequency spectrum
join <- merge(JPT[which(JPT$JPT_AF <= 1),],
              NAG[which(NAG$NAG_AF <= 1),],
              by = c('Chr', 'Pos'), all = TRUE)
join[is.na(join)] <- 0 #missing positions are frequency of 0

@@

hwe <- merge(NAG_hwe[, c('V1', 'V2')],
             JPT_hwe[, c('V1', 'V2')],
             by = c('V1', 'V2'),
             all = TRUE)
#remove sites from AF that don't meet HWE
join <- join[!paste(join$Chr, join$Pos)
             %in% c(paste(hwe$V1, hwe$V2)), ]

join <- join[, c('Chr', 'Pos', 'JPT_MAF', 'NAG_MAF')]
#make numeric
join <- sapply(join, as.numeric)
join <- join[complete.cases(join),]

@@

#save to file
saveRDS(join,
        paste0(DataPath,
               'jointAF.rds'))
}
```

@@

#get positions of random SNPs as control
set.seed(123)
random.pos <- NotSig[sample(nrow(NotSig), 
                            nrow(sig.pos)),
                     c('Chr','Pos')]
write.table(random.pos, 
            file = paste0(DataPath,
                        'random.pos.txt'), 
            col.names = FALSE, 
            row.names = FALSE, 
            quote = FALSE, 
            sep = '\t')
rm(NotSig) #save memory
```
extract context for snps of interest
```{bash}
#Using GenomeWide.AncestralContext.txt
#To get mutation context, use $path/AncestralContext.py
#compare first two columns in both files, extract matches
path='/Users/luke/Documents/QualityPaper/Misc/'
if [ ! -f $path/random.pos.context.txt ]; then
  echo "random pos context"
  awk 'NR==FNR{a[$1,$2];next} ($1,$2) in a' \
  $path/random.pos.txt \
  $path/GenomeWide.AncestralContext.txt \
  > $path/random.pos.context.txt
fi

@@

##Data Quality Over Time
```{r}
BasPath = '/Users/luke/genomes/bas_file/1000GenomesBAS/'

@@

metaData
#save to file
saveRDS(metaData,
            paste0(DataPath,
                   'metaData.rds'))
```
##Overlap of Significant SNPs from Single Population test
```{r}
phase1 <- c('YRI','LWK','CHB','CHD','JPT','CEU','LWK')
#read single population regressions
Reg <- fread(paste0(DataPath,
                    'AllRegressionsOver10.txt'), 
             col.names = c('Chr','Pos',
                           'Population','deviance'), 
             colClasses = c('numeric', 'numeric', 
                            'character', 'numeric'))
#Get P-Values and significant sites
Reg$plog10 <- -log10(pchisq(Reg$deviance, 1, 
                            lower.tail = FALSE))
Reg<-Reg[which(Reg$plog10 >= 6),]
#Count number of populations where a variant is significant
NumPop <- as.data.frame(table(Reg$Chr,Reg$Pos))
NumPop <- NumPop[which(NumPop$Freq>1),]
names(NumPop) <- c('Chr','Pos', 'SigHits')
NumPop$Chr<-as.numeric(as.character(NumPop$Chr)) 
NumPop$Pos<-as.numeric(as.character(NumPop$Pos))

@@

#save to file
saveRDS(overLap,
            paste0(DataPath,
                   'overLap.rds'))
```
## Compare Han Resequenced to 1kGP
##Single Population Test
```{r}
#load 1kgp 
kgp1= paste0(DataPath,
             '1kGP_genomeWide_SINGLEPOP_83.genotypes.txt')
kgp1.n=fread(paste0(DataPath,
                    '1kGP_genomeWide_SINGLEPOP_83.header'))
gt1<-unique(fread(kgp1,
                  col.names = colnames(kgp1.n)))
gt1.m<-melt(gt1, 
            id.vars = c('#CHROM','POS','ID'))
colnames(gt1.m) = c('CHROM','POS',
                    'ID','sample','kGT')

@@

##Full Model
```{r}
#read Genotypes
han1=fread(paste0(DataPath,
                  'SignificantSNPs_Han90_0.01_genotype.txt'))
HanNames=fread(paste0(DataPath,
                      'SignificantSNPs_Han90_COLNAMES.txt'),
               header=F)
names(han1) = as.character(unlist(HanNames))
#melt to long format
han = melt(han1, id=c('#CHROM','POS','ID'))
#fix chromosome names
han$CHROM = as.numeric(as.character(substring(han$'#CHROM',4)))
han$'#CHROM' <- NULL

@@

#save to file
saveRDS(INDEL,
            paste0(DataPath,
                   'INDEL_HanKgp.rds'))
#save to file
saveRDS(SNP,
            paste0(DataPath,
                   'SNP_HanKgp.rds'))

SNP
INDEL
```
##Compare Over and Under Qual 30
```{r}
#read Genotypes
GT = fread(paste0(DataPath,
                   'SigVar_Pop_0.01_rsID.genotype.txt'))
ColNames = read.table(paste0(DataPath,
                   'ColNames_rsID.txt'), 
                   header=F)

names(GT) = as.character(unlist(ColNames))
names(GT)[4] = 'rsID'
GT$ID <- as.numeric(as.character(GT$ID))

Sig.rsID <- fread(paste0(DataPath,
                         'SignificantVariants.txt'), 
                  col.names = c('Chr','Pos','rsID', 
                                'ID', 'dev', 'log10_p'))
##Compare over and under 30

mGT = melt(unique(GT), 
           id = c('Chr','Pos',
                  'ID','rsID'))
mGT = merge(unique(bas_dt[,
                          c('sample',
                            'Population',
                            'avg_qual')]), 
            mGT, 
            by.x='sample',
            by.y='variable')

mGT = merge(mGT, 
            Sig.rsID, 
            by=c('Chr','Pos',
                 'ID','rsID'))

over30 = mGT[which(mGT$avg_qual > 30),]

under30 = mGT[which(mGT$avg_qual < 30),]

o30AF <- 
		over30 %>% 
		group_by(Chr,Pos,
		         ID,rsID) %>% 
		summarize(oAF = sum(value))

u30AF <- 
		under30 %>% 
		group_by(Chr,Pos,
		         ID,rsID) %>% 
		summarize(uAF = sum(value))

AF30 <- merge(u30AF, 
              o30AF, 
              by=c('Chr','Pos',
                   'ID','rsID'))

#save to file
saveRDS(AF30,
            paste0(DataPath,
                   'AF30.rds'))
```
